\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage{verbatim}
\usepackage{tikz}
\usepackage{hyperref}
\setbeamertemplate{footline}[frame number]
\title{Quick SAS Tips}
\author{\texorpdfstring{Heide Jackson \newline\url{heidej@umd.edu}}{Author}}
\institute{University of Maryland Population Research Center}

\date{August 2019}

\begin{document}
\maketitle
\begin{frame}{High Level Things to Know}
\begin{itemize}
\item SAS is a statistical software package well suited to handling large data sets
\item Federal data sets are commonly stored as SAS data files (these end in the extension sas7bdat)
\item SAS can handle multiple data sets simultaneously and keep them in its remote memory.
\item SAS is not case sensitive and has extensions for using other languages particularly SQL and Python.
\end{itemize}
\end{frame}


\begin{frame}{SAS Versions}
\begin{itemize}
\item SAS 9,4 is available for download for University of Maryland affiliates.
\item SAS can be accessed online via SAS Studio
\end{itemize}
\end{frame}

\begin{frame}[fragile]{The Basics of Doing Anything in SAS}{Loading in Data}
\begin{itemize}
\item Most data is loaded, modified and saved within a data statement.  This corresponds to: 
\begin{verbatim}
data new; /*Saves new data*/
set existing; /*Loads existing data set*/
run; /*Runs or compiles the statement*/
\end{verbatim}
\end{itemize}
\end{frame}


\begin{frame}[fragile]{The Basics of Doing Anything in SAS}{Analyzing Data}
\begin{itemize}
\item Once the data is loaded, analysis can be run using the following general structure.
\begin{verbatim}
proc somecommand data=existing;
intermediate statements;
run;
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Modifying a Data Set}{Subsetting Data}
\begin{itemize}
    \item Variables can be created, modified and deleted within data statements.
    \item If a variable is dropped or kept in the set statement, this controls what is loaded in.
    \item If a variable is dropped or kept in the data statement, this controls what is saved in the new data set
    \item data can be subset with a simple if statement
    \item An example:
    \begin{small}
    \begin{verbatim}
    /*Keeps these variables in the data set being made*/
    data (keep=id var1 var2); 
    /*Drops variable from the data being loaded in*/
    set (drop=var3); 
    if var1=1; /*Keeps observations only when var1=1*/
    /*IMPORTANT changes aren't implemented until run*/
    run; 
    \end{verbatim}
    \end{small}
\end{itemize}

\end{frame}
\begin{frame}[fragile]{Modifying a Data Set}{Creating or Modifying Variables}
\begin{itemize}
    \item Variables can be created using an equal statement.
    \item Variables can be modified with if then statements.
    \item Some examples of the language used are shown on the next slide.
    \end{itemize}
    \end{frame}
    
    \begin{frame}[fragile]{Modifying a Data Set}{Creating or Modifying Variables -- Examples}
    \begin{verbatim}
    data something;
    set somethingelse;
    /*Creates a variable var1, always equal to 1*/
    var1=1; 
    /Creates a variable var1, always equal to var 2*/
    var1=var2; 
    /*If var1=1 then var2 is modified to be equal to 2*/
    if var1=1 then var2=2; 
    /*Modifies var2 if both conditions are met*/
    if var1=1 and var3=1 then var2=2; 
    ; /Modified var2 if either condition is met*/
    if var1 IN (1:5) or var3=1 then var2=2
    /*IMPORTANT changes aren't implemented until run*/
    run; 
    \end{verbatim}
\end{frame}


\begin{frame}[fragile]{Finding and Working with Data}
\begin{itemize}
\item By default SAS looks for data and stores data within a temporary working directory.  This can be found in SAS's explorer window.
\item Data sets also exist in a directory internal to sas called sashelp.
\item Users can tell SAS to look at external directories with a libname statement.
\begin{verbatim}
libname somename "/specified/full/location";
\end{verbatim}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Working with Outside Data Locations}
\begin{itemize}
\item Once a libname is specified, data can be loaded in and saved in other locations.
\item Data can be saved by adding such a location to the data step.

\item Here's a hypothetical example:
\begin{verbatim}
libname save "/specified/full/location";
libname load "/specified/full/location2";

data save.data;
set load.read;
run;
\end{verbatim}
\end{itemize}
\end{frame}





\begin{frame}[fragile]{Exporting Data}
\begin{itemize}
    \item Data can be easily exported to other programs and format.
    \item Two formats frequently used are xlsx (excel) and dta (stata).
    \begin{verbatim}
    proc export data=somesasdata filename="\full 
    location\
    dataname.extension" dbms=extension;
    /*Example of common extensions 
    are .xlsx or .dta*/
    run;
        \end{verbatim}
    \end{itemize}
    \end{frame}


\begin{frame}{General Suggestions for Getting Started}
\begin{itemize}

    \item The commands shown here can be entered via SAS's program editor.
        \item The program editor will be the script that allows users to create and reproduce data and results.
        \item Always use the log file to confirm that code was executed correctly.
        \item Unless data needs to be shared, there isn't a need to save it to a physical location.  The working directory works.
        \item It is good practice to use a different name for data in the data step versus the set step--this prevents data from being overwritten.

\end{itemize}
\end{frame}



\end{document}
